---
title: "Option Valuation"
date: "March 5, 2016"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this chapter we'll discuss the basis of option valuation in complete markets.  To help understand this method, we'll value options by the binomial method.  Lastely, we'll discuss valuation in the Black-Scholes world.

## Complete Markets

The idea of a complete market is that we can replicate any security in the market through trading in other securities.  So say a market with a risk-free bond, a stock, and an option on that stock is compelte.  This means, we can create an option through trading in the bond and the stock.  The actual option is a 'redundant' security.  

If the amrket is complete, then the value of the option is simply the cost of recreating it with the bond and the stock.  So we don't value the option *per se*, we create something that is exactly the same as the option and note its cost to create.

Could the option price be any different than the value of recreating it?  No.  If it were this would afford an arbitrage.  If the optoin cost more than the cost to create it, we could sell the option, create it, and net the difference.  A similar arbitrage obtains if the replication cost is above the option price. 

We can use a bond and stock to replicate an option:
```{r echo = FALSE, warning = FALSE}
require(visNetwork, quietly = TRUE)
# minimal example
nodes <- data.frame(id = 1:3,
  label = c("Bond", "Stock", "Option"),
  shape = c("circle", "circle", "circle"),
  group = c("A", "A", "B")
)
edges <- data.frame(from = c(1,2,1), to = c(3,3,2),
                    arrows = c("to", "to", ""))
visNetwork(nodes, edges, width = "100%")

```

Or an option and stock to replicate a bond:
```{r echo = FALSE, warning = FALSE}
require(visNetwork, quietly = TRUE)
# minimal example
nodes <- data.frame(id = 1:3,
  label = c("Option", "Stock",  "Bond"),
  shape = c("circle", "circle", "circle"),
  group = c("A", "A", "B")
)
edges <- data.frame(from = c(1,2,1), to = c(3,3,2),
                    arrows = c("to", "to", ""))
visNetwork(nodes, edges, width = "100%")
```
