---
title: "Option Valuation"
date: "March 5, 2016"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this chapter we'll discuss the basis of option valuation in complete markets.  To help understand this method, we'll value options by the binomial method.  Lastely, we'll discuss valuation in the Black-Scholes world.

## Complete Markets

The idea of a complete market is that we can replicate any security in the market through trading in other securities.  So say a market with a risk-free bond, a stock, and an option on that stock is compelte.  This means, we can create an option through trading in the bond and the stock.  The actual option is a 'redundant' security.  

If the amrket is complete, then the value of the option is simply the cost of recreating it with the bond and the stock.  So we don't value the option *per se*, we create something that is exactly the same as the option and note its cost to create.

Could the option price be any different than the value of recreating it?  No.  If it were this would afford an arbitrage.  If the option cost more than the cost to create it, we could sell the option, create it, and net the difference.  A similar arbitrage obtains if the replication cost is above the option price. 


## Binomial Option Valuation

We start option valuation with the one-period binomial method.  If you understand this, you will understand everything of importance with respect to option valuation.  Beyond this, is simply applying mathematics to make the argument work in more realistic settings.  The underlying economic argument is the same.  

Say we have the following securities:
- one-year zero-coupon risk-free bond with 5% yield.
- stock priced at $100 today.  The stock will either be $110 or $90 in one year.  
- option on the stock struck at $105 with a 1 year tenor.

Note the option will either be worth $5 or $0 in 1 year.  

### First Approach: Using the bond and stock to replicate the option.

```{r echo = FALSE, warning = FALSE}
require(visNetwork, quietly = TRUE)
# minimal example
nodes <- data.frame(id = 1:3,
  label = c("Bond", "Stock", "Option"),
  shape = c("circle", "circle", "circle"),
  group = c("A", "A", "B")
)
edges <- data.frame(from = c(1,2,1), to = c(3,3,2),
                    arrows = c("to", "to", ""))
visNetwork(nodes, edges, width = "100%")

```

We will create a portfolio of the bond and stock which pays the same as the option in both states of the world.

#### Replicating portfolio:
- sell a bond with a $90 face value, which gives us $\frac{90}{(1 + 0.05)} = \$85.71$ today.
- buy one share of the stock for $100.

This portfolio costs us $\$100 - \$85.71 = \$14.29$ to create.  In one year the portfolio will pay \$20 in the up-state (the stock is $110), and \$0 in the down state (the stock is \$90).

The option pays \$5 in the up-state and \$0 in the down-state.  The replicating portfolio we created is exactly 4 options, and therefore the cost to create the portfolio must be 4 times the cost of a single option.  This gives us the option's value: $\frac{\$14.29}{4} = \$3.57$.

If the market price is not \$3.57, can you create an arbitrage?

### Second Approach: Using the option and stock to replicate the bond.

```{r echo = FALSE, warning = FALSE}
require(visNetwork, quietly = TRUE)
# minimal example
nodes <- data.frame(id = 1:3,
  label = c("Option", "Stock",  "Bond"),
  shape = c("circle", "circle", "circle"),
  group = c("A", "A", "B")
)
edges <- data.frame(from = c(1,2,1), to = c(3,3,2),
                    arrows = c("to", "to", ""))
visNetwork(nodes, edges, width = "100%")
```

We will create a portfolio of the option and stock which pays the same as the bond in both states of the world.

#### Replicating portfolio:
- sell 1 one-year call option with a $105 strike (the same that we are valuing).
- buy $\frac{1}{4}$ share of the stock for $25.

At time 0 we have a cash flow of $C - $25, where $C is the price of the call option.  At time 1, the replicating portfolio will pay:
- up-state: $\$\frac{110}{4} - \$5 = \$22.50$
- down-state: $\$\frac{90}{4} - \$0 = \$22.50$

So no matter what happens, the replicating portfolio will pay \$22.50 in one year.  This is a risk-free zero-coupon bond, and therefore the price today is $\$\frac{22.50}{(1+0.05)} = \$21.43$.  So we have:

$\$C - \$25 = -\$21.43 \Rightarrow \$C = \$3.57$

This is exactly the same call option value we got using the first method.  

Again, if the market price is not \$3.57, can you create an arbitrage?

### Binomial Option Value Calculator

```{r echo=FALSE}
inputPanel({
 sliderInput(inputId = "stockUp", label = "Stock Price in Up-State", min = 101, max = 120, step = 1, value = 110),
 sliderInput(inputId = "stockDn", label = "Stock Price in Down-State", min = 80, max = 99, step = 1, value = 90),
 sliderInput(inputId = "rf", label = "Risk-Free Rate", min = .01, max = .05, step = .01, value = .01)
})
  
  if((input$rf + 1) * 100 > input$stockUp) {
    renderPrint({cat("Why bother with the option when you can have free money!")})
  } else {
    renderPlot({
      ## calculate value of option using first approach ----
      
      
    })
  }
  
```


#### The Importance of Zeros

Note above we used zero-coupon bonds. Zeros are important for derivative valuation because we know exactly what our return will be when we hold them (unlike with coupon bonds). 

## Beyond the Binomial Method

The binomial method highlights the important features of option valuation in complete markets.

#. We value options by creating a portfolio which replicates the option, and rely on arbitrage to ensure the prices of the replicating portfolio and the option are the same.
#. Because we replicate a risk-free bond (in the second case), we discount that bond with the risk-free rate. 
#. When delta hedging, we remove the effect of changes in the stock's price on your replicating portolio.  We therefore, don't need to know the expected return on the stock (and so we never have to specify the probability of the up and down states).


### Black-Scholes-Merton (BSM)

BSM is the seminal model in option pricing, because it is the first to use the complete markets (replication) method of finding the option's value.  Prior to BSM, option pricing models required the user to input an expected rate of return on the asset underlying the option.  Because we have no certain way to get the expected retun, we could only state the option's value in terms of this unknown parameter.  

The paper by Black and Scholes (1973) was the first to realize you 

Interestingly, the BSM option price is the limit of the binomial method as the time increment approaches zero.  However, BSM was derived prior to the binomial method.  


#### A Little Bit of History

Prior to BSM the  
